---
title: "Interchangeability Al Mansour 5 Readers"
author: "Johann Jacoby"
date: "`r format(Sys.time(), format='%Y%m%d %H%M')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libsnfct, echo=FALSE, message=FALSE, warning=FALSE}
### packages used here
pkgs2use <- list(
    c("readxl","1.3.1")
  , c("geepack","1.3-2")
  , c("multgee","1.8.0")
  , c("boot","1.3-28")
  , c("DescTools","0.99.44")
  , c("vcd","1.4-8")
  , c("ggplot2","3.3.5")
  , c("rcompanion","2.4.1")
)
### install packages if needed and load them
for (i in pkgs2use){
  if (!require(i[1], character.only = TRUE)){install.packages(i[1], character.only=TRUE)}
  library(i[1], character.only = TRUE)
}
```

```{r settings, echo=FALSE}
cilevel <- 95 # confidence level for cis (in percent)
digr <- 4 # default rounding digit number
digrkappa <- 2 # default rounding digit number for kappa/kendalls tau/W
nboot <- 10000 # bootstrap resamples
doboot <- TRUE # turn off bootstrapping by setting FALSE

```

```{r dataprep, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
filename <- "DL_Spine_interchangeability_Wirbelkörperebene_R1-R5.xlsx" # file name with extension without any path information, within quotes
path <- "" # path to directory where <filename> resides, in quotes, without trailing slash, system specific slash hard coded
setwd(path)

rawd <- as.data.frame(read_excel(filename, sheet = 1))
d<-rawd
# streamline var names
names(d) <- gsub("^Marrow","marrow", names(d))
names(d) <- gsub("^foram_","foramen_", names(d))

# recode variables with errors or too little variance for interchangeability analysis

# recode Modic variables into dichotomous because of tiny numbers of Type 1 and Type 3 in sample
sapply(grep("^Modic", names(d)), function(x) d[, x] <<- ifelse(is.na(d[,x]), NA, ifelse(d[,x]==0,0,1)))
# recode Pfirr variables, so that 0 and 1 are collapsed to category 0 and all others to 1, too few diagnostically positive cases
sapply(grep("^Pfirr", names(d)), function(x) d[, x] <<- ifelse(is.na(d[,x]), NA, ifelse(d[,x]>1,1,0)))
# recode Schm variables, so that values of 2 (only 4 cases by Reader 4: 2 under TSE, 2 under DL) are coded as 1
sapply(grep("^Schm_", names(d)), function(x) d[, x] <<- ifelse(is.na(d[,x]), NA, ifelse(d[,x]>0,1,0)))
# recode Flav variables, so that values of 2 (only 2 cases by Reader 4: 1 under TSE, 1 under DL) are coded as 1
sapply(grep("^Flav_", names(d)), function(x) d[, x] <<- ifelse(is.na(d[,x]), NA, ifelse(d[,x]>0,1,0)))
# recode Facet variables, correct error value "2" to "0"
sapply(grep("^Facet", names(d)), function(x) d[, x] <<- ifelse(is.na(d[,x]), NA, ifelse(d[,x]==2,0,d[,x])))
# recode foramen variables, so that cases = 0 are category 0 and all others (>0) are collapsed to 1, too few diagnostically positive cases
sapply(grep("^foramen", names(d)), function(x) d[, x] <<- ifelse(is.na(d[,x]), NA, ifelse(d[,x]>0,1,0)))


vars <- unique(sapply(names(d)[-1], function(x) strsplit(split="_", x)[[1]][1]))
raters <- unique(sapply(names(d)[-1], function(x) strsplit(split="_", x)[[1]][3]))

d$PatientID <- sapply(d$Patient, function(x) paste0(strsplit(split="_", x)[[1]][2:3], collapse=""))
d$region <- sapply(d$Patient, function(x) strsplit(split="_", x)[[1]][2])
d$PatientItem <- sapply(d$Patient, function(x) strsplit(split="_", x)[[1]][4])

stacks <- list()

for (v in vars){
  for (r1 in 1:length(raters)){
    for (r2 in 1:length(raters)){
      if (r2 > r1) { # within standard method
        stacks[[length(stacks)+1]] <- cbind(
          d[,c("PatientID", "region", "PatientItem")]
          , j1=d[,paste0(v,"_TSE_R",r1)], j2=d[,paste0(v,"_TSE_R",r2)]
          , agree=as.numeric(ifelse(d[,paste0(v,"_TSE_R",r1)]==d[,paste0(v,"_TSE_R",r2)],1,0))
          , agreeOrd=sapply(abs(d[,paste0(v,"_TSE_R",r1)]-d[,paste0(v,"_TSE_R",r2)]), function(x) ifelse(x == 0, 1, ifelse(x==1, .5, 0)))
        , comp="TSEwithin", compclass="within", r1=r1, r2=r2, m1="TSE", m2="TSE", var=v
        )
      }
      ## between methods for all rater pairs
      stacks[[length(stacks)+1]] <- cbind(
        d[,c("PatientID", "region", "PatientItem")]
        , j1=d[,paste0(v,"_TSE_R",r1)], j2=d[,paste0(v,"_DL_R",r2)]
        , agree=as.numeric(ifelse(d[,paste0(v,"_TSE_R",r1)]==d[,paste0(v,"_DL_R",r2)],1,0))
        , agreeOrd=sapply(abs(d[,paste0(v,"_TSE_R",r1)]-d[,paste0(v,"_DL_R",r2)]), function(x) ifelse(x==0, 1, ifelse(x==1, .5, 0)))
      , comp="TSE.DL", compclass="between", r1=r1, r2=r2, m1="TSE", m2="DL", var=v
      )
    }
  }
}
LL<-do.call(rbind, stacks)
LL$compclass.refwithin<-relevel(factor(LL$compclass, levels=c("between","within")), ref="within")
LL$PatientID<-factor(LL$PatientID)
LL<-LL[order(LL$PatientID),]

dput(x=LL, file = "LLdat.dput")


dlonglong <- reshape(d, direction="long", varying=2:ncol(d), timevar="variable", v.names="judgement", times=names(d)[2:ncol(d)])
dlonglong[, c("parameter","method","reader")] <- do.call(rbind,sapply(dlonglong$variable, function(x) strsplit(split="_",x)))
dlonglong$parameter <- factor(dlonglong$parameter, levels=vars, labels=vars)
descplot<-ggplot(data=dlonglong, aes(x=as.numeric(judgement), fill=reader)) + geom_histogram(position="dodge") + facet_wrap(. ~ parameter, nrow = 2)
descplot


geeboot <- function(theData, w){
  bssample <- theData[w,]
  mm<-geeglm(agree ~ 1 + compclass.refwithin, id=PatientID, data=na.omit(bssample), family=binomial(link="logit"),corstr="exchangeable")$coefficients
  unname((exp(mm[1]+mm[2]) / (1+exp(mm[1]+mm[2]))) - (exp(mm[1])/(1+exp(mm[1]))))
}

resulttbl <- as.data.frame(do.call(rbind, lapply(vars, function(v) {
  cat(paste0("resulttbl",v))
  dichvar <- ifelse(nrow(table(unname(unlist(d[,grep(paste0("^",v,".+"), names(d))])))) == 2,TRUE,FALSE)
  datahere <- subset(LL, var==v)
  datahere <- datahere[order(datahere$PatientID),]
  intWN <- intB <- eff <- NA
  if (dichvar) {
    model <- geeglm(agree ~ compclass.refwithin, id=PatientID, data=datahere, family=binomial , corstr="exchangeable")
    intWN <- (exp(model$coefficients[1])/(1 + exp(model$coefficients[1])))
    intBW <- (exp(model$coefficients[1] + model$coefficients[2])/(1+exp(model$coefficients[1] + model$coefficients[2])))
    if(doboot) { bo <- boot(data=datahere, R=nboot, statistic=geeboot, sim="ordinary") }
  } else {
    model <- ordLORgee(formula=agreeOrd ~ compclass.refwithin, id=PatientID, data=datahere, LORstr="uniform")
    rvs <- c(1,.5,0)
    res <- model$fitted.values[regData$method==0,][1,]
    intWN <- res[1]*rvs[1]+res[2]*rvs[2]+res[3]**rvs[3]
    res2 <- model$fitted.values[regData$method==1,][1,]
    intBW <- res2[1]*rvs[1]+res2[2]*rvs[2]+res2[3]**rvs[3]
    if (doboot) { bo <- boot(data=datahere, R=nboot, statistic=multgeeboot, parallel="snow") }
  }
  eff <- unname(intBW-intWN)
  perc.bs.ci <- c(0,0)
  if(doboot) { perc.bs.ci <- boot.ci(bo, type="perc")$percent[c(4,5)] }
  
 #calculate and prepare fleiss' kappas and cohen's kappas
 dkar <- d[,grep(paste0(v, "_TSE"), names(d))]
 dkrr <- d[,grep(paste0(v, "_(TSE|DL)"), names(d))]
 dkra <- as.data.frame(apply(do.call(rbind, lapply(1:length(raters), function(x) as.matrix(cbind(d[, c(paste0(v,"_",c("TSE","DL"),"_R",x))])))),2,function(x) as.numeric(x)))

 kappa.ar <- KappaM(x=dkar, method = "Fleiss", conf.level=cilevel/100)
 kappa.rr <- KappaM(x=dkrr, method = "Fleiss", conf.level=cilevel/100)
 kappa.ra <- Kappa(x=table(dkra), weights="Equal-Spacing")

 individualkappas<-c()
 individualkappa.cis<-c()
   for (rr in 1:length(raters)){
     subdata<-d[,grep(paste0(v, "_(TSE|DL)_R",rr), names(d))]
     if(!all(dim(table(subdata))==2)) { 
      individualkappas[rr] <- individualkappa.cis[rr] <- NA
     } else {
     assign(x=paste0("kappa.ra.",rr), value=Kappa(x=table(subdata), weights="Equal-Spacing"))
     individualkappas[rr] <- get(paste0("kappa.ra.",rr))$Unweighted["value"]
     if(!dichvar){individualkappas[rr] <- get(paste0("kappa.ra.",rr))$Weighted["value"]}
     individualkappa.cis[rr] <- paste0("[",paste0(round(digits=digr, confint(get(paste0("kappa.ra.",rr)),level = cilevel/100)["Unweighted",]), collapse=";"),"]")
     if (!dichvar) {
       individualkappa.cis[rr] <- paste0("[",paste0(round(digits=digr, confint(get(paste0("kappa.ra.",rr)),level = cilevel/100)["Weighted",]), collapse=";"),"]")
     }
    }
   }
 kappa.ra.mean <- mean(individualkappas, na.rm=TRUE)

   #prepare and calculate kendall's ws and kendall's taus
   kendallw.ar <- kendallW(x=dkar, ci=TRUE, conf=cilevel/100)
   kendallw.rr <- kendallW(x=dkrr, ci=TRUE, conf=cilevel/100)
   kendalltaub.ra <- KendallTauB(table(dkra), conf.level=cilevel/100)
   kendalltaub.rasp <- sprintf(paste0("%.",digrkappa,"f"), round(digits=digrkappa, kendalltaub.ra))
   
   # prepare and calculate findings table  
casesz<-data.frame(do.call(rbind, lapply(names(d)[grep(paste0(v,"_TSE_R"), names(d))], function(x) cbind(d[, x], d[, gsub("TSE","DL",x)]))))
tbz<-table(factor(levels=c("11","10","01","00"), apply(casesz, 1, function(x) paste0(ifelse(x[1]>0,1,0),ifelse(x[2]>0,1,0)))))
casesb<-data.frame(
    tse=apply(d[, grep(paste0(v,"_TSE_R"), names(d))], 1, function(y) ifelse(sum(y==1, na.rm=TRUE)>0,1,0))
  , dl=apply(d[, grep(paste0(v,"_DL_R"), names(d))], 1, function(y) ifelse(sum(y==1, na.rm=TRUE)>0,1,0))
)
tb<-table(factor(levels=c("11","10","01","00"),apply(casesb, 1, function(x) paste0(x[1],x[2]))))
 
  findings11z<-as.numeric(tbz["11"])
  findings01z<-as.numeric(tbz["01"])
  findings10z<-as.numeric(tbz["10"])
  findings00z<-as.numeric(tbz["00"])
  findingstotalz<-sum(findings11z,findings10z,findings01z)
  findingsNz<-sum(findings11z,findings10z,findings01z,findings00z)
  findingstotalzp<-(findingstotalz/findingsNz)*100
  findings11z.p<-(findings11z/findingstotalz)*100
  findings01z.p<-(findings01z/findingstotalz)*100
  findings10z.p<-(findings10z/findingstotalz)*100
  findings00z.p<-(findings00z/findingstotalz)*100
  findingstestpz <- mcnemar.test(matrix(ncol=2, byrow=TRUE, c(findings11z,findings10z,findings01z,findings00z)))$p.value
   
  findings11<-as.numeric(tb["11"])
  findings01<-as.numeric(tb["01"])
  findings10<-as.numeric(tb["10"])
  findings00<-as.numeric(tb["00"])
  findingstotal<-sum(findings11,findings10,findings01, na.rm=TRUE)
  findingsN<-sum(findings11,findings10,findings01,findings00, na.rm=TRUE)  
  findingstotalp<-(findingstotal/findingsN)*100
  findings11.p<-(findings11/findingstotal)*100
  findings01.p<-(findings01/findingstotal)*100
  findings10.p<-(findings10/findingstotal)*100
  findings00.p<-(findings00/findingstotal)*100
  findingstestp <- mcnemar.test(matrix(ncol=2, byrow=TRUE, c(findings11,findings10,findings01,findings00)))$p.value

  ## draw everything in list to be rbinded into a large table with all results per variable
list(
    Variable=v
    , bootstraps=nboot
    , bootstrapdone=doboot
    ### Intercepts = agreement proportion estimates
    , InterceptWN=unname(intWN)
    , InterceptBW=unname(intBW)
    ### IEI & CIs
    , IEI=eff
    , IEI.text=sprintf(paste0("%.",digr,"f"), round(digits=digr,eff))
    , CI.IEI.ll=perc.bs.ci[1]
    , CI.IEI.ul=perc.bs.ci[2]
    , CI.IEI.text=paste0("[",paste0(sprintf(paste0("%.",digr,"f"), round(digits=digr, perc.bs.ci)), collapse=";"),"]")
    ### kappas
    , kappa.ar.with.CI=
     paste0(sprintf(paste0("%.",digrkappa,"f"), kappa.ar[1])," [",paste0(sprintf(paste0("%.",digrkappa,"f"),  round(digits=digrkappa, kappa.ar[c(2,3)])), collapse=";"),"]")
    , kappa.rr.with.CI=
      paste0(sprintf(paste0("%.",digrkappa,"f"), round(digits=digrkappa, kappa.rr[1]))," [",paste0(sprintf(paste0("%.",digrkappa,"f"),  round(digits=digrkappa, kappa.rr[c(2,3)])), collapse=";"),"]")
    , kappa.ra.with.CI=
        paste0(sprintf(paste0("%.",digrkappa, "f"), round(digits=digrkappa, kappa.ra$Unweighted["value"]))," [",paste0(sprintf(paste0("%.",digrkappa,"f"), round(digits=digrkappa, confint(kappa.ra)["Unweighted",])), collapse=";"),"]")
    ### Kendall W / tau
    , kendallsw.intraprotocol.interreader = sprintf(paste0("%.",digrkappa,"f"), round(digits=digrkappa, as.numeric(kendallw.ar[1])))
    , kendallsw.intraprotocol.interreader.with.CI = paste0(sprintf(paste0("%.",digrkappa,"f"), round(digits=digrkappa, as.numeric(kendallw.ar[1])))," [",paste0(sprintf(paste0("%.",digrkappa,"f"), round(digits=digrkappa, as.numeric(kendallw.ar[2:3]))), collapse=";"),"]")
    , kendallsw.interprotocol.interreader = sprintf(paste0("%.",digrkappa,"f"), round(digits=digrkappa, as.numeric(kendallw.rr[1])))
    , kendallsw.interprotocol.interreader.with.CI = paste0(sprintf(paste0("%.",digrkappa,"f"), round(digits=digrkappa, as.numeric(kendallw.rr[1])))," [",paste0(sprintf(paste0("%.",digrkappa,"f"), round(digits=digrkappa, as.numeric(kendallw.rr[2:3]))), collapse=";"),"]")
    , kendallsw.interprotocol.intrareader.with.CI = paste0(kendalltaub.rasp[1]," [", kendalltaub.rasp[2], ";", kendalltaub.rasp[3], "]")
    ### findings
    , findings11z=findings11z
    , findings01z=findings01z
    , findings10z=findings10z
    , findings00z=findings00z
    , findingstotalz=findingstotalz
    , findingstotalzp=findingstotalzp
    , findingsNz=findingsNz
    , findings11z.p=findings11z.p
    , findings01z.p=findings01z.p
    , findings10z.p=findings10z.p
    , findings00z.p=findings00z.p
    , findings11=findings11
    , findings01=findings01
    , findings10=findings10
    , findings00=findings00
    , findingstotal=findingstotal
    , findingstotalp=findingstotalp
    , findingsN=findingsN
    , findings11.p=findings11.p
    , findings01.p=findings01.p
    , findings10.p=findings10.p
    , findings00.p=findings00.p
    , findingstestp=findingstestp
    , findingstestpz=findingstestpz
)
})))

row.names(resulttbl) <- unlist(resulttbl$Variable)
resulttbl<- resulttbl[unlist(unname(sapply(vars, function(x) if(x %in% row.names(resulttbl)) {x} ))),]


write.table(apply(resulttbl, 2, function(x) as.character(x)), file = paste0(format(Sys.time(), format="%Y%m%d.%H%M%S"),"_resulttable.tsv"), sep = "\t", row.names = TRUE, quote=FALSE)


#prepare table for exchangeability indices
resulttblprintnamesiei <- list(
  Variable="Variable"
  , InterceptWN="Agreements TSL~TSL"
  , InterceptBW="Agreements TSL~DL"
  , IEI.text="IEI"
  , CI.IEI.text=paste0(cilevel,"%CI Bootstrapped Percentile of IEI")
)

#prepare table kappas/kendalls w and tau
resulttblprintnameskappa <- list(
    kappa.ar.with.CI="Fleiss' &kappa; Intra method Inter reader"
  , kappa.rr.with.CI="Fleiss' &kappa; Inter method Inter reader"
  , kappa.ra.with.CI="Cohen's &kappa; Inter method Intra reader"
  , kendallsw.intraprotocol.interreader.with.CI="Kendall's W Intra method Inter reader"
  , kendallsw.interprotocol.interreader.with.CI="Kendall's W Inter method Inter reader"  
  , kendallsw.interprotocol.intrareader.with.CI="Kendall's &tau; Inter method Intra reader"  
)


resulttblprintnamesfindingsz <- list(
    findingsNz="N"
  , findingstotalz="Total number of findings"
  , findingstotalzp="Total number of findings % "  
  , findings11z="Findings both methods"
  , findings11z.p="Findings both methods %"
  , findings10z="Findings by standard method TSE, but not DL"
  , findings10z.p="Findings TSE, not DL %"
  , findings01z="Findings by new method DL, but not TSE"
  , findings01z.p="Findings DL, not TSE %"
  , findingstestpz="p (McNemar test)"
)


resulttblprintnamesfindings <- list(
    findingsN="N"
  , findingstotal="Total number of findings"
  , findingstotalp="Total number of findings %"  
  , findings11="Findings both methods"
  , findings11.p="Findings both methods %"
  , findings10="Findings by standard method TSE, but not DL"
  , findings10.p="Findings TSE, not DL %"
  , findings01="Findings by new method DL, but not TSE"
  , findings01.p="Findings DL, not TSE %"
  , findingstestp="p (McNemar test)"
)


resulttblprintiei <- resulttbl[,names(resulttblprintnamesiei)]
resulttblprintkappa <- resulttbl[,names(resulttblprintnameskappa)]
resulttblprintfindingsz <- resulttbl[,names(resulttblprintnamesfindingsz)]
resulttblprintfindings <- resulttbl[,names(resulttblprintnamesfindings)]

names(resulttblprintiei) <- sapply(names(resulttblprintiei), function(x) resulttblprintnamesiei[[x]])
names(resulttblprintkappa) <- sapply(names(resulttblprintkappa), function(x) resulttblprintnameskappa[[x]])
names(resulttblprintfindingsz) <- sapply(names(resulttblprintfindingsz), function(x) resulttblprintnamesfindingsz[[x]])
names(resulttblprintfindings) <- sapply(names(resulttblprintfindings), function(x) resulttblprintnamesfindings[[x]])

sapply(c(3,5,7,9), function(x) {
  resulttblprintfindingsz[,x] <<- sprintf(paste0("%.",digr,"f"), round(digits=digr, as.numeric(resulttblprintfindingsz[,x])))
  resulttblprintfindings[,x] <<- sprintf(paste0("%.",digr,"f"), round(digits=digr, as.numeric(resulttblprintfindings[,x])))
})



```  

# Table Exchangeability Index  

`r knitr::kable(resulttblprintiei)`
CIs are percentile-based bootstrapped (`r sprintf("%.0f", nboot)` resamples).


# Table Agreements  

This table contains various inter/intra method and inter/intra reader agreement statistics.  

**Intramethod interreader agreement and intermethod interreader agreement**  
Fleiss' &kappa; (with `r length(raters)` readers) and Kendall's w.


**Intermethod intrareader agreements** 
Cohen's &kappa; and Kendall's &tau; for simple method agreement TSE vs. DL (with pairs of judgements of individual readers stacked as if they were independent)

`r knitr::kable(resulttblprintkappa)`

# Table Findings  




## alternative: at least one reader (two judgments per vertebra: TSE and DL, readers combined within TSE & within DL)

Finding in TSE: when at least one reader rated present per vertebra in all TSE ratings  
Finding in DL: when at least one reader rated present per vertebra in all DL ratings  
1 Rating pair per vertebra (finding using TSE - finding using DL)  

`r knitr::kable(resulttblprintfindings, digits=2)`  
